\begin{center} \bf \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\[-1cm] \end{center}

### 2021-05-25 main.tex
first draft to CH

\begin{abstract}
Repeated games have been the centerpiece when studying the evolution of human
cooperation. Most of the existing models work with idealized assumptions.
Individuals are assumed to interact with a representative sample of the
population, and they remember all interactions they participate in. In real
life, however, individuals do not always remember all interactions they had with
everyone in their population. Instead, they rather recall the last interaction
they had.

In this work we explore the effect on the evolution of cooperation if
individuals only remember a minimum of social information. We present results
for a series of simulations were we vary the benefit of cooperation and the
strength of selection. We present results for the most commonly used classes of
social dilemmas, and we show that even though cooperation can evolve when
individuals only use a minimum of information, the evolving cooperation rates
are typically lower than in the classical scenario.
\end{abstract}

@@

One of the most important applications of evolutionary game theory is the
evolution of cooperation. Why is it that some individuals choose to help others,
increasing their payoff, at the expense of decreasing one’s own? In evolutionary
game theory individuals are not required to be rational, instead they adapt
strategies based on mutation and exploration. Strategies are more likely to
spread if they have a high fitness either because the individuals who adopt them
have more offspring, or because they are imitated more often~\cite{Wu2015}. The
fitness of a strategy is not constant but depends on the composition of the
population. Individuals interact based on their strategies with other members of
the population and the payoffs they yield are translated into fitness.

It is commonly assumed that individuals compute their fitness after interacting
with a representative sample of the population, and remembering all the
interactions they participated in~\cite{Nowak2006}. Thus, they imply that
individuals have a perfect memory. However, when modelling how individuals make
decisions in each turn they are assumed to have very limited memory. To be
precise, most of the works in the literature focus on naive subjects who can
only choose from a restricted set of strategies~\cite{Nowak1992tit}, or who do
not remember anything beyond the outcome of the very last round~\cite{Baek2016}.
Note that there are a few notable exceptions~\cite{Hauert1997, Stewart2016}.
The perfect memory assumption is not only unrealistic but it also creates
this curious inconsistency.

This has led us to question the robustness of our understanding of cooperation.
In this work we explore whether cooperation can evolve if individuals compute
their fitness based on a minimum of social information. Though we are not the
first to question the assumptions of estimating fitness~\cite{Roca2006}, we are
the first to explore the effect of payoff memory. %ToDo add more information

We first consider two extreme scenarios, the classical scenario of the expected
payoffs and the alternative scenario where individuals update their strategies
only based on the very last payoff they obtained. We observe that individuals
with limited memory tend to adopt less generous strategies and they achieve less
cooperation than in the classical scenario. We obtain similar results when we
consider that individuals update their strategies based on more social
information. More specifically, up to the last two payoffs they obtained when
interacting with up to two different members of the population.

@@

We consider a population of \(N\) players~\footnote{The terms ``player'' and
``individual'' are used interchangeably here.} where \(N\) is even and mutations
are sufficiently rare. Therefor, at any point in time there are at most two
different strategies present in the population; a \textit{resident} strategy and
a \textit{mutant} strategy. To describe how strategies spread we use a pairwise
comparison process~\cite{Traulsen2006}.
Each step of the evolutionary process consists of two
stages, a game stage and an updating stage.

In the game stage each individual is randomly matched with some other individual
in the population. They engage in a match where each subsequent turn
occurs with a fixed probability $\delta$. At each turn players can choose to
either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
depend on both their decisions. If both players cooperate they receive the
reward payoff \(R\), whereas if both defect they receive the punishment payoff
\(P\). If one cooperates but the other defects, the defector receives the
temptation \(T\) payoff, whereas the cooperator receives the sucker's
payoff \(S\). We denote the feasible payoff of each turn as \(\mathcal{U} =
\{R, S, T, P\}\).
We assume that individuals use \textit{reactive strategies} to make
decisions in each turn. Reactive strategies are a set of memory-one strategies
that only take into account the previous action of the opponent. They can be
written explicitly as a vector \(\in \R_{3}\), more specifically, a reactive
strategy \(s\) is given by \(s=(y, p, q)\) where \(y\) is the probability that
the strategy opens with a cooperation and \(p, q\) are the probabilities that
the strategy cooperates given that the opponent cooperated and defected
equivalently.

In the updating stage, two players are randomly drawn from the population, a
`learner' and a `role model'. Given that the learner's payoff $u_L\!\in\!
\mathcal{U}$ and that the role model's payoff $u_{RL}\!\in\! \mathcal{U}$.
The learner adopts the role model's strategy based on the Fermi
distribution function,

@@

where $\beta\!\ge\!0$ is the intensity of selection. The updating payoffs of the
learner and the role model are conventionally correspond to their expected
payoffs. The expected payoff of a strategy is the mean payoff of the strategy of
engaging in repeated games with all other members of the population. The expected payoffs
assume perfect memory. In this work we define a new a set of updating
payoffs. We reefer to these as the limited memory payoffs, and we compare them
to the classical scenario.

The evolutionary step is repeated until either the mutant strategy goes
extinct, or until it fixes in the population. If the mutant fixes in the
population then the mutant strategy becomes the new resident strategy. After
either outcome we introduce a new mutant strategy uniformly chosen from all
reactive strategies at random, and we set the number of mutants to $1$. This
process of mutation and fixation/extinction is then iterated many times.
More details on our methodology are found in Appendix~\ref{appendix:methods}.

In order to account for the effect of the updating payoffs approaches, we
explore the cooperation rate within the resident population over multiple
generations. In the next section we present a series of simulation results.
To account for the various types of social behaviour we also present
results on multiple social dilemmas.

@@

In this section we explore the case where the updating payoffs are based on the
last round payoff achieved against another member of the population and we compare
this to the classical scenario of the expected payoffs. We assume that each pair
of players interacts in a donation game. The donation game is a special case of
the prisoner's dilemma. Each player can choose to cooperate by providing a
benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
Thus, the feasible payoffs in each round are \(T=b, R=b-c, S=-c, P=0\).

@@

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}\vspace{-3cm}
    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
    \caption{{\bf Evolutionary dynamics under expected payoffs and last round with one interaction payoffs.} 
    We have run two simulations of the evolutionary process described in
    section~\ref{section:model} for $T\!=\!10^7$ time steps. For each time step,
    we have recorded the current resident population ($y,p,q$). Since simulations
    are run for a relatively high continuation probability of $\delta\!=\!0.999$, we
    do not report the players' initial cooperation probability $y$. The graphs show
    how often the resident population chooses each combination ($p,q$) of
    conditional cooperation probabilities in the subsequent rounds. ({\bf A}) If
    players update based on their expected payoffs, the resident population
    typically applies a strategy for which $p\!\approx\!1$ and
    $q\!\le\!1\!-\!c/b\!=\!0.9$. ({\bf B})
    When players update their strategies based on their realized payoffs in the last
    round, there are two different predominant behaviors. The resident population
    either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
    conditional cooperators. In the latter case, the maximum level of $q$ consistent
    with stable cooperation is somewhat smaller compared to the expected-payoff
    setting, $q\!<\!0.5$. The cooperation rate within the resident population
    (averaged over all games and over all time steps) is close to 100\%. Parameters:
    $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
    \label{fig:expected_and_stochastic_for_donation}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=.8\textwidth]{static/expected_for_beta.pdf}
    \includegraphics[width=.8\textwidth]{static/stochastic_for_beta.pdf}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=.8\textwidth]{static/cooperation_rate_over_b.pdf}
  \end{subfigure}
  \caption{{\bf The evolution of cooperation for different benefit values.} 
  We vary the benefit of defection $b$. In all cases, expected payoffs appear to
  overestimate the average cooperation rate the population achieves. ({\bf A})
  the probabilities \(p, q\) for resident population over \(10^7\) time steps
  for each benefit value. ({\bf B}) The cooperation rate within the resident population
  (averaged over all games and over all time steps) over the benefit.
  Unless
  explicitly varied, the parameters of the simulation are $N\!=\!100$,
  $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run
  for $T\!=\!5\times 10^7$ time steps for each parameter
  combination.}\label{fig:cooperation_rate_over_benefit}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{.65\textwidth}
    \centering
    \includegraphics[width=.8\textwidth]{static/expected_for_selection_strenght.pdf}
    \includegraphics[width=.8\textwidth]{static/stochastic_for_selection_strenght.pdf}
  \end{subfigure}%
  \begin{subfigure}{.35\textwidth}
    \vspace{.3cm}
    \includegraphics[width=.8\textwidth]{static/cooperation_rate_over_betas.pdf}
  \end{subfigure}
\caption{{\bf The evolution of cooperation for different selection strength values.}
We vary the selection strength $\beta$. In all cases, stochastic payoff
evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
probabilities \(p, q\) for resident population over \(10^7\) time steps for each
\(\beta\) value. ({\bf B}) The cooperation rate within the resident population
(averaged over all games and over all time steps) over \(\beta\).
Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
$b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
$T\!=\!5\times 10^7$ time steps for each parameter
combination.}\label{fig:cooperation_rate_over_betas}
\end{figure}

@@

In the previous section we gained insights into the effects of the updating
payoffs, and into how parameters such as the benefit and the strength of
selection can intensify them. We investigated these effects by using the
donation game. In order to broaden our understanding of the updating payoffs on
different forms of possible human interactions we extend our approach to other
\(2 \times 2\) symmetric games. More specifically, we apply our analysis to four
different classes of games as given by Table~\ref{table:social_dilemmas}.

\begin{table}[!htbp]
  \begin{center}
    \resizebox{.4\textwidth}{!}{
    \renewcommand{\arraystretch}{1.5}
      \begin{tabular}{ccc}
  \toprule
  & \textbf{social dilemmas} & \textbf{preference ordering} \\
  \midrule
(i) & harmony  & \(R > T > S > P\) \\
(ii) & stag hunt  & \(R > T > P > S\) \\
(iii) & prisoner dilemma  & \(T > R > P > S\) \\
(iv) & snowdrift  & \(T > R > S > P\) \\
  \bottomrule
      \end{tabular}}
  \end{center}
\caption{\textbf{Social dilemmas and preference ordering}. The four classes of
social dilemmas we explore in this work.}
\label{table:social_dilemmas}
\end{table}

We compare the results of the evolutionary process described in
section~\ref{section:model} when the updating payoffs are based on the expected
payoffs, and on the last round payoffs for all the four possible classes of
games. The results of the simulations are given in
Figures~\ref{fig:expected_two_by_two} and~\ref{fig:last_round_two_by_two}
respectively.

In Figure~\ref{fig:expected_two_by_two} each sub figure represents a run of
evolutionary process for a different set of values for \(\mathcal{U}\). Without
loss of generality we set \(R=1\) and \(P=0\)~\cite{Martinez2012, Roca2009}, and
we vary the values of the temptation payoff \(T\) (across the \(x\)-axis) and of
the sucker's payoff \(S\) (across the \(y\)-axis). Starting at the upper left
corner and proceeding clockwise the quadrants correspond to the harmony, the
snowdrift, the prisoner's dilemma and the stag hunt games.

The harmony game represents are social situation without conflict. It is in the
best interest for both players to cooperate as \(R > T\) and \(S > P\).
Figure~\ref{fig:expected_two_by_two} confirms this. We observe that for most
harmony game runs the resident population overwhelmingly applies a strategy for
which \(p\) and \(p\) are \(\approx 1\). In the case of \(T = R = 1\) we observe
that the resident strategies become slightly less generous. Against a cooperating
strategy defecting and cooperation yield the same result. The lower \(q-\)values
could indicate that the resident strategy is less generous to defectors to avoid
being invaded.

The snowdrift game describes a situation similar to that of the prisoner's
dilemma. Cooperation results in a benefit to the opposing player but entailing a
cost to the cooperator. However, in the snowdrift game individuals obtain
immediate direct benefits from the cooperative acts which leads to \(S > P\). In
the snowdrift game more cooperation emerges compared to the prisoner's dilemma.
Defection is never a resident strategy and for the shame values of temptation
the overall \(q-values\) are higher. It can be seen that the \(p-values\) are
lower as individuals have free room to defect after receiving a cooperation.

The last class of games we present are for the stag hunt game. In the stag hunt
both players benefit for mutual cooperation and defection, however, \(R > P\)
and thus the resident strategies with the highest fitness are the cooperative
ones. The results for the stag hunt class are similar to those of the harmony
game.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
  \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.} 
  We have run several simulations of the evolutionary process described in
  section~\ref{section:model} for $T\!=\!10^7$ time steps. The graphs show how
  often the resident population chooses each combination ($p,q$) of conditional
  cooperation probabilities in the subsequent rounds. We vary the temptation
  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
  across the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$.}
  \label{fig:expected_two_by_two}
\end{figure}

The conclusions regarding the evolved behaviour for the various classes of
games remain the same when the last round payoffs are considered,
Figure~\ref{fig:last_round_two_by_two}. However, there is variation in these
populations. This variation as an effect affects the resident within cooperative
rates.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
  \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
  We have run several simulations of the evolutionary process described in
  section~\ref{section:model} for $T\!=\!10^7$ time steps. The graphs show how
  often the resident population chooses each combination ($p,q$) of conditional
  cooperation probabilities in the subsequent rounds. We vary the temptation
  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
  across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$.}
  \label{fig:last_round_two_by_two}
\end{figure}

The cooperation rates for each of the case of the social dilemmas we have used
are given in Figure~\ref{fig:cooperation_rates_two_by_two_part_one}. In the case
of the harmony game the differences are very small, even in the case of \(T=1\)
where the biggest difference occurs it still remain less than 10\%. In the case
of the stag hunt game the difference increases in the case of \(T=1\). The
biggest variation in the results are for the class of the prisoner's dilemma.
For values of \(S < -0.4\) the are significantly different, in the case of the
last round payoffs cooperation almost never is a valid option. Overall the last
round payoffs are less cooperative, supporting the results of the previous
section. There are only two cases which that is not true and both cases are in
the snowdrift class. For the values of \(T>1.8\) and \(S=0.4, 0.8\) cooperation
is estimated higher. In the last round payoffs a Tit For Tat like behaviour is
more robust that in the case of the expected payoffs.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
  \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
  \caption{\textbf{Cooperation rates over various social dilemmas.} ({\bf A}) If
  players update based on their expected payoffs. ({\bf B}) If
  players update based on their last round payoffs. We vary the temptation
  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
  across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$.}\label{fig:cooperation_rates_two_by_two_part_one}
\end{figure}

So far we have explored the difference between the expected payoffs and the last
round payoffs. In order to explore further the effect of limited memory we allow
individuals to remember more. We consider the cases where individuals remember
up two interactions, and up to the last two rounds. In total we present results
for three more cases. These are the cases of the last round two rounds with
another member of the population, last round with two members of the population,
and last round two rounds with two members of the population. Similarly to the
previous examples we have run the evolutionary process for a large number of
step for each of the social dilemmas. The results are found in the
Appendix~\ref{appendix:further_simulation_results}.

>>>>

\begin{abstract}
Repeated games are vastly used in evolutionary game theory to explain
cooperation in a variety of environments. Although existing models have shaped
our understanding of human cooperation, the often work with idealized
assumptions. In an evolutionary process, individuals imitate other individuals
of the population based on their fitness. It is commonly assumed that an
individual computes their fitness after interacting with a representative sample
of the population, and remembering all the interactions they participated in. In
real life, we do not always remember all our interactions, instead we rather
recall our most recent ones.

Here, we introduce a framework that allows individuals to estimate their fitness
based on a minimum of social information. We explore the difference between the
classical framework and ours using computer simulations. We present results for
the most commonly used classes of symmetric \(2 \times 2\) games. The
simulations show that, in the prisoner's dilemma individuals with limited memory
tend to adopt less generous strategies and they achieve less cooperation than in
the classical scenario. In contrast, in the stag-hunt and the harmony game, the
impact of memory is less striking. Here individuals with limited memory perform
nearly as well as individuals with full memory. Finally, we observe that the
snowdrift game is the only game for which cooperation can be underestimated in
the classical scenario.
\end{abstract}

@@

One of the most important applications of evolutionary game theory is the
evolution of cooperation. Why is it that some individuals choose to help others,
increasing their payoff, at the expense of decreasing one's own? In evolutionary
game theory individuals are not required to be rational, instead they adapt
strategies based on mutation and exploration. Strategies are more likely to
spread if they have a high fitness either because the individuals who adopt them
have more offspring, or because they are imitated more often~\cite{Wu2015}. The
fitness of a strategy is not constant but depends on the composition of the
population. Individuals interact based on their strategies with other members of
the population and the payoffs they yield are translated into fitness.

It is commonly assumed that individuals compute their fitness after interacting
with a representative sample of the population, and remembering all the
interactions they participated in~\cite{Nowak2006, nowak2004emergence,
Hilbe2013, Hilbe2018,Reiter2018, garcia2018no, Nowak1992tit, Baek2016}. This
implies that individuals have a perfect memory. However, when modelling how
individuals make decisions in each turn they are assumed to have very limited
memory. To be precise, most of the studies in the literature focus on naive
subjects who can only choose from a restricted set of
strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
outcome of the very last round~\cite{Baek2016}. Note that there are a few
notable exceptions~\cite{Hauert1997, Stewart2016}.

Studies that make use of this classical framework report that cooperation can
substantially evolve. However, the inconsistency in the memory size of players
leads us to question the robustness of our understanding of cooperation. To this end, we
propose a framework in which individuals, similar to the decisions at each turn,
estimate their fitness based on a minimum of information. Though we are not the
first to question the assumptions of estimating fitness~\cite{Roca2006}, we are
the first to explore the effect of memory.

We first consider two extreme scenarios, the classical scenario
and the alternative scenario where individuals update their strategies
only based on the very last payoff they obtained. We observe that individuals 
with limited memory tend to adopt less generous strategies and they achieve less
cooperation when interacting in a prisoner's dilemma. We obtain similar results when we
consider that individuals update their strategies based on more
information. More specifically, up to the last two payoffs they obtained when
interacting with up to two different members of the population. We extend our
approach to the rest of the symmetric \(2 \times 2\) games.

@@

We consider a population of \(N\) players~\footnote{The terms ``player'' and
``individual'' are used interchangeably here.} where \(N\) is even and mutations
are sufficiently rare. Therefore, at any point in time there are at most two
different strategies present in the population; a \textit{resident} strategy and
a \textit{mutant} strategy. To describe how strategies spread we use a pairwise
comparison process~\cite{Traulsen2006}.
Each step of the evolutionary process consists of two
stages, a game stage and an updating stage.

In the game stage each individual is randomly matched with some other individual
in the population. They engage in a match where each subsequent turn
occurs with a fixed probability $\delta$. At each turn players choose independently to
either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
depend on both their decisions. If both players cooperate they receive the
reward payoff \(R\), whereas if both defect they receive the punishment payoff
\(P\). If one cooperates but the other defects, the defector receives the
temptation payoff \(T\), whereas the cooperator receives the sucker's
payoff \(S\). We denote the feasible payoff of each turn as \(\mathcal{U} =
\{R, S, T, P\}\).
We assume that individuals use \textit{reactive strategies} to make
decisions in each turn. Reactive strategies are a set of memory-one strategies
that only take into account the previous action of the opponent. They can be
written explicitly as a vector in \(\R_{3}\), more specifically, a reactive
strategy \(s\) is given by \(s=(y, p, q)\). The parameter \(y\) is the probability that
the strategy opens with a cooperation and \(p, q\) are the probabilities that
the strategy cooperates given that the opponent cooperated and defected
equivalently.

In the updating stage, two players are randomly drawn from the population, a
`learner' and a `role model'. Given the learner's payoff $u_L\!\in\!
\mathcal{U}$ and the role model's payoff $u_{RL}\!\in\! \mathcal{U}$,
the learner adopts the role model's strategy with probability,

@@

where $\beta\!\ge\!0$ is the strength of selection. For small values of $\beta$
the imitation probability is independent of the strategies of the involved
players. As the value of $\beta$ increases, the more likely it is that the
learner adopts only strategies that yield a higher payoff.
Conventionally the updating payoffs of the learner and the role model are based
on their expected payoffs. A player's expected payoff is the mean payoff the
player yields after engaging in matches of multiple turns with each member of
the population. At each match a player bases their next turn decision only on
the previous action of the opponent, however, the same player bases their
expected payoffs on the outcomes of all their matches. Thus, a player is assumed
to have limited and perfect memory at the same time. We propose a new a set of
updating payoffs where it is also assumed that a player has limited memory.
We referee to these as the limited memory payoffs.

The evolutionary step is repeated until either the mutant strategy goes
extinct, or until it fixes in the population. If the mutant fixes in the
population then the mutant strategy becomes the new resident strategy. After
either outcome we introduce a new mutant strategy uniformly chosen from all
reactive strategies at random, and we set the number of mutants to $1$. This
process of mutation and fixation/extinction is then iterated many times.

In order to account for the effect of the updating payoffs we simulate the
evolutionary process and record which strategies the players adopt over time
based on (i) the expected payoffs (ii) the limited memory payoffs. We compare
the difference in the cooperation rate within the resident population for the
two approaches. To account for the various types of social behaviour we also
present results on multiple social dilemmas.

@@

In this section we explore the case where the updating payoffs are based on the
last round payoff achieved against another member of the population and we compare
this to the classical scenario of the expected payoffs. We assume that each pair
of players interacts in a donation game. The donation game is a special case of
the prisoner's dilemma. Each player can choose to cooperate by providing a
benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
Thus, the feasible payoffs in each round are \(\mathcal{U} = \{b-c, -c, b, 0\}\).

@@

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}\vspace{-3cm}
    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
    \caption{{\bf Evolutionary dynamics under expected payoffs and last round with one interaction payoffs.} 
    We have run two simulations of the evolutionary process described in
    section~\ref{section:model} for $t\!=\!10^7$ time steps. For each time step,
    we have recorded the current resident population ($y,p,q$). Since simulations
    are run for a relatively high continuation probability of $\delta\!=\!0.999$, we
    do not report the players' initial cooperation probability $y$. The graphs show
    how often the resident population chooses each combination ($p,q$) of
    conditional cooperation probabilities in the subsequent rounds. ({\bf A}) If
    players update based on their expected payoffs, the resident population
    typically applies a strategy for which $p\!\approx\!1$ and
    $q\!\le\!1\!-\!c/b\!=\!0.9$. ({\bf B})
    When players update their strategies based on their realized payoffs in the last
    round, there are two different predominant behaviors. The resident population
    either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
    conditional cooperators. In the latter case, the maximum level of $q$ consistent
    with stable cooperation is somewhat smaller compared to the expected-payoff
    setting, $q\!<\!0.5$. The cooperation rate within the resident population
    (averaged over all games and over all time steps) is close to 100\%. Parameters:
    $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
    \label{fig:expected_and_stochastic_for_donation}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
    \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
  \end{subfigure}
  \caption{{\bf The evolution of cooperation for different benefit values.} 
  We vary the benefit of defection $b$. In all cases, expected payoffs appear to
  overestimate the average cooperation rate the population achieves. ({\bf A})
  the probabilities \(p, q\) for resident population over \(10^7\) time steps
  for each benefit value. ({\bf B}) The cooperation rate within the resident population
  (averaged over all games and over all time steps) over the benefit.
  Unless
  explicitly varied, the parameters of the simulation are $N\!=\!100$,
  $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run
  for $t\!=\!5\times 10^7$ time steps for each parameter
  combination.}\label{fig:cooperation_rate_over_benefit}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{.65\textwidth}
    \centering
    \includegraphics[width=\textwidth]{static/expected_for_selection_strenght.pdf}
    \includegraphics[width=\textwidth]{static/stochastic_for_selection_strenght.pdf}
  \end{subfigure}%
  \begin{subfigure}{.35\textwidth}
    \vspace{.3cm}
    \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
  \end{subfigure}
\caption{{\bf The evolution of cooperation for different selection strength values.}
We vary the selection strength $\beta$. In all cases, stochastic payoff
evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
probabilities \(p, q\) for resident population over \(10^7\) time steps for each
\(\beta\) value. ({\bf B}) The cooperation rate within the resident population
(averaged over all games and over all time steps) over \(\beta\).
Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
$b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
$t\!=\!5\times 10^7$ time steps for each parameter
combination.}\label{fig:cooperation_rate_over_betas}
\end{figure}

@@

In the previous section we gained insights into the effects of the updating
payoffs, and into how parameters such as the benefit and the strength of
selection can intensify them. We investigated these effects by using the
donation game. In order to broaden our understanding of the updating payoffs on
different forms of possible human interactions we extend our approach to all
\(2 \times 2\) symmetric games. More specifically, we apply our analysis to four
different classes of games; the harmony, the snowdrift, the prisoner's dilemma
and the stag-hunt games.

We compare the results of the evolutionary process when the updating payoffs are based on the expected
payoffs, and on the last round payoffs for all the four possible classes of
games (Figures~\ref{fig:expected_two_by_two}
and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
each sub figure represents a run of evolutionary process for a different set of
values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
\(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
the stag-hunt games.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
  \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
  We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
  the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
  \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
  corresponds to the harmony game. The preference ordering for the harmony game
  is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T > 1\)
  corresponds to the snowdrift game. The preference ordering for the snowdrift game
  is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where \(S < 0\) and \(T \leq 1\)
  corresponds to the stag-hunt game. The preference ordering for the stag-hunt game
  is \(R > T > P > S\). ({\bf D}) The bottom right quadrant where \(S < 0\) and \(T > 1\)
  corresponds to the prisoner's dilemma game. The preference ordering for the prisoner's dilemma game
  is \(T > R > P > S\).
  Unless explicitly varied, the parameters of the simulation
  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
  \label{fig:expected_two_by_two}
\end{figure}

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
  \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
  We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
  the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
  \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
  corresponds to the harmony game. The preference ordering for the harmony game
  is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T > 1\)
  corresponds to the snowdrift game. The preference ordering for the snowdrift game
  is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where \(S < 0\) and \(T \leq 1\)
  corresponds to the stag-hunt game. The preference ordering for the stag-hunt game
  is \(R > T > P > S\). ({\bf D}) The bottom right quadrant where \(S < 0\) and \(T > 1\)
  corresponds to the prisoner's dilemma game. The preference ordering for the prisoner's dilemma game
  is \(T > R > P > S\).
  Unless explicitly varied, the parameters of the simulation
  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
  \label{fig:last_round_two_by_two}
\end{figure}

The harmony game describes an interaction without conflict. In the harmony game
it is in the best interest for both players to cooperate, and in an evolutionary
process cooperators prevail. We observe that for most harmony game runs the
resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
describes a situation similar to that of the prisoner's dilemma; cooperation
results in a benefit to the opposing player but entailing a cost to the
cooperator. However, in the snowdrift game individuals obtain immediate direct
benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
game more cooperation emerges compared to the prisoner's dilemma. Defection is
never a resident strategy and for the same values of temptation the overall
\(q-\) values are higher. The last class of games we present are for the stag-hunt
game. In the stag-hunt game there is an equilibrium in which both players
cooperate as well as one in which both defect. For small values of the
temptation and the sucker's payoffs we observe that cooperation prevails similar
to the harmony game. However, in the case where the temptation and the reward
payoffs are equal the resident population either consists of defectors or
cooperators.

Although the patterns of the evolved populations remain similar if we consider
that players update their strategies based on their last round payoff
(Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
between the two approaches. To better understand these dissimilarities, we
explore the cooperation rates for each of the cases of the social dilemmas we
have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).

In the harmony game the cooperation rates for both approaches are similar. The
biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
are two instances for which the cooperation rates are higher for the limited
memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
\(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
overestimated when a minimum of information is considered. Players are slightly
more likely to cooperate after a cooperation and to forgive after being at the
receiving end of a defection. For the rest of the cases the classical scenario
overestimates the average cooperation rate, supporting the results of the
previous section. There is a substantial difference in the cooperation rates
for the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation
almost never emerges if players imitate based only on the last round payoff.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
  \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
  \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
  players update based on their expected payoffs. ({\bf B}) If
  players update based on their last round payoffs. We vary the temptation
  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
  across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
\end{figure}
